Online Multiclass Linear Classification with Bandit Feedback

### 2018-11-13 arxiv-version/main.tex
Add game protocol and a reference

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{natbib}

@@

We study the online multiclass classification problem with bandit feedback. The
problem was initially proposed by \cite{Kakade-Shalev-Shwartz-Tewari-2008}. The
corresponding problem with full information feedback is solved by the classical
\textsc{Multiclass Perceptron} algorithm and its variants (see e.g.
\cite{Crammer-Dekel-Keshet-Shalev-Shwartz-Singer-2006}). The problem is quite
natural and potential applications include online advertising and online content
optimization.

In this paper, we focus on the special case when the examples are \emph{linearly
separable with a margin}. In the full-information feedback setting, it is well
know that if the examples lie in the ball of radius $R$ centered at the origin
and is linearly separable with a margin $\gamma$ then the \textsc{Multiclass
Perceptron} algorithm makes at most $(R/\gamma)^2$ mistakes. This result is very
satisfactory since the upper bound on the number of mistakes is
information-theoretically optimal and at the same time the \textsc{Multiclass
Perceptron} algorithm has low time and memory complexity.

The bandit feedback setting, however, is much more challenging. We start by
dissecting the notion of linear separability. We define two notions of linear
separability for multiclass classification. the ``standard'' linear separability
and \emph{strong linear separability}. For binary classification, these two
definitions are identical. However, they differ for multiclass classification.
For multiclass classification with $K$ clases the standard notion of linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside the class lie in the complement of the
intersection of the halfspaces. Strong linear separability means that examples
from each class can be separated from the remaining examples by a \emph{single}
hyperplane. We will give a more precise definition
in Section~\ref{section:notions-of-linear-separability}.

We provide a simple and efficient algorithm for the bandit feedback setting if
the examples are strongly linearly separable. The algorithm runs $K$ copies a
\textsc{Binary Perceptron} algorithm.

@@

Finally, we study two questions related to the computational and
information-theoretic hardness of the problem. Any algorithm for the bandit
setting collects information in the form of so called \emph{strongly labeled}
and \emph{weakly labeled} examples. Strongly-labeled examples are those for
which we know the class label. Weakly labeled example is an example for which we
know that class label can be anything except for a particular one class. First,
we show the offline problem of determining if a linear multiclass separator
exists for a set of strongly- and weakly-labeled examples is NP-complete.
Second, we show that any algorithm that uses only strongly-labeled examples and
ignores weakly labeled examples makes at least $\Omega(???)$ mistakes.

>>>>

% LaTeX packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{natbib}

% Math commands
\newcommand{\indicator}[1]{\mathbf{1}\left[#1 \right]}

@@

We study the online multiclass classification problem with bandit feedback. The
protocol of the problem is stated below. The problem was initially proposed by
\cite{Kakade-Shalev-Shwartz-Tewari-2008}. The corresponding problem with full
information feedback is solved by the classical \textsc{Multiclass Perceptron}
algorithm and its variants (see e.g.
\cite{Crammer-Dekel-Keshet-Shalev-Shwartz-Singer-2006}). The problem is quite
natural and potential applications include online advertising and online content
optimization where the feedback corresponds to a user's click.

\begin{algorithm}[h]
\caption{Online multiclass classification with bandit feedback
\label{algorithm:game-protocol}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$}
\FOR{$t=1,2,\dots$}
\STATE{Receive feature vector $x_t$}
\STATE{Environment \emph{secretly} chooses correct class label $y_t \in \{1,2,\dots,K\}$}
\STATE{Predict class label $\widehat y_t \in \{1,2,\dots,K\}$}
\STATE{Receive feedback $z_t \in \{0,1\}$ where $z_t = \begin{cases}
1 & \text{if $\widehat y_t = y_t$,} \\
0 & \text{if $\widehat y_t \neq y_t$.}
\end{cases}$
}
\ENDFOR
}
\end{algorithmic}
\end{algorithm}

In this paper, we focus on the special case when the examples are \emph{linearly
separable with a margin}. In the full-information feedback setting, it is well
know that if the examples have norm at most $R$ and are linearly separable with
a margin $\gamma$ then the \textsc{Multiclass Perceptron} algorithm makes at
most $(R/\gamma)^2$ mistakes. This result is very satisfactory since the upper
bound on the number of mistakes is information-theoretically optimal and at the
same time the \textsc{Multiclass Perceptron} algorithm has low time and memory
complexity.

The bandit feedback setting, however, is much more challenging. We start by
dissecting the notion of linear separability. We define two notions of linear
separability for multiclass classification: the standard linear separability
and \emph{strong linear separability}. For binary classification, these two
definitions are identical. However, they differ for multiclass classification.
For multiclass classification with $K$ classes the standard notion of linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside the class lie in the complement of the
intersection of the halfspaces. Strong linear separability means that examples
from each class can be separated from the remaining examples by a \emph{single}
hyperplane. We state the definitions precisely
in Section~\ref{section:notions-of-linear-separability}.

If the examples are strongly linearly separable, there is a simple and efficient
algorithm for the bandit feedback setting that makes at most $?/\gamma^2$
mistakes. The algorithm runs $K$ copies of the \textsc{Binary Perceptron}
algorithm.

@@

Finally, we study two questions related to the computational and
information-theoretic hardness of the problem. Any algorithm for the bandit
setting collects information in the form of so called \emph{strongly labeled}
and \emph{weakly labeled} examples. Strongly-labeled examples are those for
which we know the class label. Weakly labeled example is an example for which we
know that class label can be anything except for a particular one class. First,
we show the offline problem of determining if a linear multiclass separator
exists for a set of strongly- and weakly-labeled examples is NP-complete. This
result is a simple consequence of the fact learning intersection of halfspaces
is NP-complete~\citep{Blum-Rivest-1993}. Second, we show that any algorithm that
uses only strongly-labeled examples and ignores weakly labeled examples makes at
least $\Omega(???)$ mistakes.
